---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading and Preprocessing the Data

```{r load-data}
activity <- read.csv("activity.csv", stringsAsFactors = FALSE)
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
str(activity)
```

# What is Mean Total Number of Steps Taken per Day?

```{r total-steps-day}
total_steps_per_day <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
hist(total_steps_per_day$steps,
     breaks = 20,
     main = "Total Steps per Day",
     xlab = "Total Steps",
     col = "lightblue",
     border = "black")
mean_steps <- mean(total_steps_per_day$steps)
median_steps <- median(total_steps_per_day$steps)
```

The mean total number of steps taken per day is approximately `r mean_steps`.  
The median total number of steps taken per day is approximately `r median_steps`.

# What is the Average Daily Activity Pattern?

```{r daily-pattern}
avg_steps_per_interval <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
plot(avg_steps_per_interval$interval, avg_steps_per_interval$steps, type = "l",
     main = "Average Daily Activity Pattern",
     xlab = "5-minute Interval",
     ylab = "Average Number of Steps",
     col = "blue")
max_interval <- avg_steps_per_interval$interval[which.max(avg_steps_per_interval$steps)]
```

The 5-minute interval with the maximum average number of steps is `r max_interval`.

# Imputing Missing Values

```{r imputing-missing-values}
total_missing <- sum(is.na(activity$steps))
total_missing
activity_imputed <- merge(activity, avg_steps_per_interval, by = "interval", suffixes = c("", "_mean"))
activity_imputed$steps[is.na(activity_imputed$steps)] <- activity_imputed$steps_mean[is.na(activity_imputed$steps)]
activity_imputed$steps_mean <- NULL
```
```{r plot-imputed-data}
total_steps_per_day_imputed <- aggregate(steps ~ date, data = activity_imputed, sum)
hist(total_steps_per_day_imputed$steps,
     breaks = 20,
     main = "Total Steps per Day (Imputed Data)",
     xlab = "Total Steps",
     col = "lightgreen",
     border = "black")
mean_steps_imputed <- mean(total_steps_per_day_imputed$steps)
median_steps_imputed <- median(total_steps_per_day_imputed$steps)
```

After imputing missing values:  
The mean total steps per day is `r mean_steps_imputed`.  
The median total steps per day is `r median_steps_imputed`.

# Are there Differences in Activity Patterns Between Weekdays and Weekends?

```{r weekday-weekend-patterns}
activity_imputed$day_type <- ifelse(weekdays(activity_imputed$date) %in% c("Saturday", "Sunday"),
                                    "weekend", "weekday")
activity_imputed$day_type <- factor(activity_imputed$day_type)
avg_steps_daytype <- aggregate(steps ~ interval + day_type, data = activity_imputed, mean)
```
```{r plot-weekday-weekend}
library(lattice)
xyplot(steps ~ interval | day_type, data = avg_steps_daytype, type = "l",
       layout = c(1, 2),
       xlab = "5-minute Interval",
       ylab = "Average Number of Steps",
       main = "Activity Patterns: Weekdays vs Weekends",
       col = "blue")
```

The panel plot shows that activity patterns differ between weekdays and weekends.  
On weekends, activity tends to be more evenly spread out during the day.
